---
title: "Practical Machine Learning"
author: "Khairul Ashraf"
date: "February 1, 2016"
output: html_document
---

###SYNOPSIS

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways.

###DATA

The training data for this project are available 
[here](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv).

The test data are available 
[here](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv).

### EXPLORATORY DATA ANALYSIS

```{r eval=FALSE} 

library(data.table)
library(caret)
library(randomForest)
library(foreach)
library(rpart)
library(rpart.plot)
library(corrplot)

#Loading the data
trainingRaw <- read.csv("pml-training.csv", na.strings=c("#DIV/0!"," ", "", "NA", "NAs", "NULL"))
testingRaw <- read.csv("pml-testing.csv", na.strings=c("#DIV/0!"," ", "", "NA", "NAs", "NULL"))

```


### DATA CLEANING

Removing columns with NAs, highly correlated variables and variables with 0 variance.

```{r eval=FALSE}


trainingClean <- trainingRaw[, -which(names(trainingRaw) %in% c("X", "user_name", "raw_timestamp_part_1", "raw_timestamp_part_2", "cvtd_timestamp", "new_window", "num_window"))]

#exclude columns with NAs
trainingData = cleanClean[, colSums(is.na(trainingClean)) == 0]

#exclude highly correlated variables
NZV =nearZeroVar(trainingData[sapply(trainingData, is.numeric)], saveMetrics=TRUE)
trainingNZV = trainingData[, zero_variance[, 'nzv'] == 0]

CorrMatrix <- cor(na.omit(trainingNZV[sapply(trainingNZV, is.numeric)]))

dim(CorrMatrix)
CorrMatrixFormat <- expand.grid(row = 1:52, col = 1:52)
#returns the correlation matrix in matrix format

CorrMatrixFormat$correlation <- as.vector(CorrMatrix)
highCorrelationRem <- findCorrelation(CorrMatrix, cutoff = .7, verbose = TRUE)
#exclude highly correlated variables

training <- trainingNZV[, -highCorrelationRem] 

for(i in c(8:ncol(training)-1)) {training[,i] = as.numeric(as.character(training[,i]))}

for(i in c(8:ncol(testingRaw)-1)) {testingRaw[,i] = as.numeric(as.character(testingRaw[,i]))}

```


Removing unnecessary dataset.

```{r eval=FALSE}

#drop blank column
setData <- colnames(training[colSums(is.na(training)) == 0])[-(1:7)]
modelset <- training[setData]

#clean data to build model
setData

```

### CROSS-VALIDATION

```{r eval=FALSE}

partitioned <- createDataPartition(setB$classe, p=0.6, list=FALSE )
train <- modelset[partitioned,]
test <- modelset[-partitioned,]

trainingControl <- trainControl(method="cv", 5)
trainModel <- train(classe ~ ., data=train, method="rf", trControl=control, ntree=250)
trainModel

prediction <- predict(model, testing)
confusionMatrix(test$classe, predict)

resultAccuracy <- postResample(prediction, test$classe)
resultAccuracy

result <- predict(setA, train[, -length(names(training))])
result

treeModel <- rpart(classe ~ ., data=training, method="class")
prp(treeModel) 

```

### SUBMISSION

The accuracy of the model can be seen from the confusion matrix.

The below shown script was used to obtain single text files to be uploaded to the courses web site to comply with the submission assigment,

```{r eval=FALSE}

pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

x <- evaluation_data
x <- x[feature_set[feature_set!='classe']]
answers <- predict(rf, newdata=x)

answers

pml_write_files(answers)
```

